---
title: "Final"
date: "2026-03-17"
author: "Bella Kormos"
format:
  html:
    toc: true # use this to display a table of contents
execute:
  message: false # use this to make sure messages don't show up
  warning: false # use this to make sure warnings don't show upl
---

[GitHub Repo](https://github.com/bellakormos/ENVS-193DS_homework-03)

# Homework Set Up

Reading in packages:

```{r}
#| label: packages
#| message: false
library(tidyverse) # load in all packages
library(janitor) # clean column names
library(here) # identifies computer's root folder
library(lubridate) # work with date/time data
library(rstatix) # load rstatix for cliff's delta
library(rcompanion) # also cliff's delta
library(effsize) # add this — contains cliff.delta()
```

Reading in data:

```{r}
#| label: reading in data
#| message: false
kelp <- read_csv(here("data", "kelp.csv")) # store the data and save object as "kelp" 
my_data <- read_csv(here("data", "official_personal_data.csv")) # Read in the personal data as an object named "my_data"
```

# Problem 1. Research writing

## a. Transparent statistical methods

## b. Figure needed

## c. Suggestions for rewriting

## d. Interpretation

# Problem 2. Data visualization

## a. Cleaning and summarizing

```{r}
#| label: data cleaning
#| message: false
#| warning: false

# creating new object called 'kelp_clean'
kelp_clean <- kelp %>% 
  # cleaning up column names
  clean_names() %>% 
  # filter by site 
  filter(site %in% c("IVEE", "NAPL", "CARP")) %>%
  # creating a new column called site_name, relabeling species with full name
  mutate(site_name = case_when(
         site == "NAPL" ~ "Naples",
         site == "IVEE" ~ "Isla Vista",
         site == "CARP" ~ "Carpinteria"
)) %>% 
  # convert to factor and set factor level order
  mutate(site_name = as_factor(site_name),
         site_name = fct_relevel(site_name, 
                                 "Naples", "Isla Vista", "Carpinteria")) %>% 
  group_by(site_name, year) %>% # group for summarizing
  summarize(mean_fronds = mean(fronds)) %>% # calculate mean fronds per group
  ungroup() # remove grouping structure
```

## b. Understanding the data

## c. Visualize the data

## d. Interpretation

# Problem 3. Data analysis

## a. Response variable

## b. Purpose of study

## c. Table of models

## d. Run the models

## e. Select the best model

The best model as determined by Akaike’s Information Criterion (AIC) (insert answer)

Use the predictors and the response variable to describe the model, not the model number that you assigned.

## f. Check the diagnostics

plot ( simulateResiduals(bat_mod) ) are they uniformally distributed? not normally.

## g. Visualize the model predictions

## h. Write a caption for your figure.

## i. Calculate model predictions

## j. Interpret your results

# Problem 4. Affective and exploratory visualizations

## a. Comparing visualizations

### How are the visualizations different from each other in the way you have represented your data?

In homework 2, I was comparing 2 variables at a time trying to find a correlative relationship. For my affective visualization, the data observations are represented artistically, abstractly, and it is harder to decode the data and find the statistical relationships.

### What similarities do you see between all your visualizations?

In the different categories of my data across all my visualizations, I use different colors/shapes to distinguish the variables. For my categorical visualization, café = gold, library = blue; for my affective visualization, jade plants = café, pilea plants = library; for the study duration vs tasks completed, the tasks completed are in green. There is definitely a creative, colorful component to all visualizations.

### What patterns (e.g. differences in means/counts/proportions/medians, trends through time, relationships between variables) do you see in each visualization?

In the continuous visualization, there is a general positive trend: as study duration increases, so does the tasks completed. In the categorical visualization, the medians of tasks completed between the library vs café in the boxplots are compared, and there is a higher median in library study sessions than the café sessions. In the affective visualization, all variables are presented in the form of houseplants and since all variables are represented, you can extrapolate relationships such as comparing the amount of friends with me or the busyness of the location.

### Are these different between visualizations? If so, why? If not, why not?

The relationships found are all relatively similar, and I am not interpreting the data in different ways across all visualizations.

### What kinds of feedback did you get during week 9 in workshop or from the instructors? How did you implement or try those suggestions? If you tried and kept those suggestions, explain how and why; if not, explain why not.

The feedback I got was that I could find a creative way to put a "code" of my variables so you can read the data observations on a coffee specials sign. I loved this idea, so I put a code of variables on the window as window art and on a sign outside the café in my painting.

## b. Designing an analysis

The appropriate analysis for my personal data is the Mann-Whitney U test.It is appropriate here because my response variable (tasks completed) is numeric and my predictor variable (study location) is categorical with two levels (café and library). Given the small sample size and non-normal distribution (non-centered medians in boxplot), a non-parametric test that compares medians is more appropriate than a t-test.

## c. Check any assumptions and run your analysis

### Check all assumptions

**Check 1: Random Sample** Each data point represents an independently recorded study session, and sessions were not deliberately selected to favor a location, so this assumption is met.

**Check 2: Visual**

```{r}
#| label: histogram-assumption
#| message: false
#| warning: false
my_clean_data <- my_data |> 
  clean_names() # columns names lower case with underscores
# base layer: ggplot using cleaned data-set
ggplot(data = my_clean_data,
       aes(x = tasks_completed, # tasks completed on x-axis
           fill = study_location)) + # fill bars by study location
  # first layer: histogram
  geom_histogram(binwidth = 1,
                 alpha = 0.7, # semi-transparent bars
                 color = "white") + # white outline between bars
  facet_wrap(~study_location) + # separate panel for each location
  scale_x_continuous(breaks = seq(0, 12, by = 1)) + # whole number x-axis ticks
  scale_fill_manual(values = c("Café" = "#FEBC11", # gold for café
                               "Library" = "#4169E1")) + #blue for library
  # Label axes and plot 
  labs(title = "Distribution of Tasks Completed by Study Location", 
       x = "Tasks Completed",
       y = "Number of Study sessions") +
  theme_linedraw() +
  theme(legend.position = "none",
        panel.grid.minor = element_blank(), # remove minor gridlines
        axis.line.x = element_line(color = "black"))
```
Both the café and library distributions are unimodal and right-skewed, indicating similar distributional shape, so all assumptions are met."

### Wilcoxon rank sum exact test

**Summary of stats: median and n values**

```{r}
#| label: summary stats
#| message: false
#| warning: false
my_clean_data %>% 
  group_by(study_location) |> # group by café vs library
  summarize(
    median = median(tasks_completed), # calculate median tasks completed
    n = n() # count number of study sessions per group
  )
```

**Run a Wilcoxon test**
```{r}
#| label: wilcoxon test
#| message: false
#| warning: false
wilcox.test(tasks_completed ~ study_location, # compare tasks completed by study location
            data = my_clean_data) # use cleaned personal data
```

**Calculate effect size using Cliff's Delta**
```{r}
#| label: effect-size
#| message: false
#| warning: false

cliff.delta(tasks_completed ~ study_location,
            data = my_clean_data)
```
I found a small (Cliff’s δ = -0.08) difference in median task completed between cafés (n = 5) and the library(n = 15). (Wilcoxon rank-sum test, W = 103.5, p = 0.72, ⍺ = 0.05). Tasks completed tended to be smaller in cafés (median = 3) than in the library(median = 5). 

## d. Create a visualization
```{r}
#| label: mann-whitney-visualization
#| message: false
#| warning: false
#| code-fold: true
# base layer: ggplot using cleaned data-set
ggplot(data = my_clean_data,
       aes(x = study_location, # categorical predictor on x-axis
           y = tasks_completed)) + # response variable on y-axis
  # first layer: boxplot
  geom_boxplot(aes(fill = study_location), # fill boxplot by study location
               width = 0.5,
               outliers = FALSE, # avoid outliers being shown twice
) + # boxplot width
  # second layer: add jittered points
  geom_jitter(width = 0.2, # horizontal spread
              height = 0, # no vertical jitter
              shape = 21) + # open circle points
  scale_fill_manual(values = c(
    "Café" = "#FEBC11", # gold for café
    "Library" = "#4169E1")) + # blue for library
  # add labels for axes and title plot
  labs(x = "Study Location",
       y = "Tasks Completed (n)") +  
  # edit theme, and add grid lines
  theme_minimal() + 
  theme(legend.position = "none", # remove legend
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"))
```

## e. Write a caption
**Figure 2. Distribution of tasks completed by study location across all study sessions (Café vs. Library)** The most recent observation: March 4, 2026. Each circle point represents a study session. The library boxplot (in blue) has higher median (5 tasks), and the café study sessions (in gold) shows a tighter interquartile range and smaller median (3 tasks).

## f. Write about your results
I did not find a significant difference in median tasks completed between café (n = 16, median = 3) and library (n = 15, median = 5) study sessions (Wilcoxon rank-sum test: W = 103.5, p = 0.72, α = 0.05), with a negligible effect size (Cliff's δ = -0.08). Although the library median was larger by 2 tasks, the large p-value and negligible effect size suggest this difference is not meaningful and could be due to random chance. Study location does not appear to have a significant influence on my productivity as measured by tasks completed per session. 


## g. Sharing your affective visualization

Attendance of week 10 workshop!
