---
title: "Final"
date: "2026-03-17"
author: "Bella Kormos"
format:
  html:
    toc: true # use this to display a table of contents
execute:
  message: false # use this to make sure messages don't show up
  warning: false # use this to make sure warnings don't show upl
---

Link to my github: [GitHub Repository](https://github.com/bellakormos/ENVS-193DS_homework-03)

# Homework Set Up

Reading in packages:

```{r}
#| label: packages
#| message: false
library(tidyverse) # load in all packages
library(janitor) # clean column names
library(here) # identifies computer's root folder

```

Reading in data:

```{r}
#| label: reading in data
#| message: false
kelp <- read_csv(here("data", "kelp.csv")) # store the data and save object as "kelp" 
```

# Problem 1. Research writing

## a. Transparent statistical methods

## b. Figure needed

## c. Suggestions for rewriting

## d. Interpretation


# Problem 2. Data visualization

## a. Cleaning and summarizing

```{r}
#| label: data cleaning
#| message: false
#| warning: false

# creating new object called 'kelp_clean'
kelp_clean <- kelp %>% 
  # cleaning up column names
  clean_names() %>% 
  # filter by site 
  filter(site %in% c("IVEE", "NAPL", "CARP")) %>%
  # creating a new column called site_name, relabeling species with full name
  mutate(site_name = case_when(
         site == "NAPL" ~ "Naples",
         site == "IVEE" ~ "Isla Vista",
         site == "CARP" ~ "Carpinteria"
)) %>% 
  # convert to factor and set factor level order
  mutate(site_name = as_factor(site_name),
         site_name = fct_relevel(site_name, 
                                 "Naples", "Isla Vista", "Carpinteria")) %>% 
  group_by(site_name, year) %>% # group for summarizing
  summarize(mean_fronds = mean(fronds)) %>% # calculate mean fronds per group
  ungroup() # remove grouping structure
```

## b. Understanding the data

## c. Visualize the data

## d. Interpretation


# Problem 3. Data analysis

## a. Response variable

## b. Purpose of study

## c. Table of models

## d. Run the models

## e. Select the best model

## f. Check the diagnostics

## g. Visualize the model predictions

## h. Write a caption for your figure.

## i. Calculate model predictions

## j. Interpret your results


# Problem 4. Affective and exploratory visualizations

## a. Comparing visualizations

## b. Designing an analysis

## c. Check any assumptions and run your analysis

## d. Create a visualization

## e. Write a caption

## f. Write about your results

## g. Sharing your affective visualization
Attendance of week 10 workshop!







